On the presentation of orthogonal groups over polynomial rings

### 2019-07-03 dl-paper.tex
Added remark to Lemma 3.3

\subsection{Steinberg groups}
Let $\Phi$ be a reduced and irreducible root system of rank $\geq 2$ and $R$ be a commutative ring with $1$. Recall that in this case the \emph{Steinberg group} $\St(\Phi, R)$ can be defined by means of generators $x_{\alpha}(s)$ (where $\alpha\in\Phi, s\in R$) and relations:
\begin{align}
& \phantom{[}
x_\alpha(s) \cdot x_\alpha(t) = x_\alpha(s+t),\ \alpha\in\Phi,\ s,t\in R; \label{rel:add}\\
& [x_\alpha(s), x_\beta(t)] = \prod\limits_{i,j\in\mathbb{N}}
 x_{i\alpha + j\beta}\left(N_{\alpha,\beta ij}\, s^i t^j\right),\quad \alpha,\beta\in\Phi,\ \alpha\neq\pm\beta,\ s,t\in R. \label{rel:CCF}
\end{align}
The indices $i$, $j$ appearing in the right-hand side of the above relation range over
all positive natural numbers such that $i\alpha + j\beta\in\Phi$.
The constants $N_{\alpha \beta i j}$ appearing in the right-hand side of \eqref{rel:CCF} are integers equal to $\pm 1,\pm 2,\pm 3$, they are called the {\it structure constants} of the Chevalley group $G_{sc}(\Phi, R)$. There is a degree of freedom in the choice of signs of the structure constants. Several different methods have been proposed for choosing signs of these constants, see e.\,g.~\cite{V00}, \cite[\S~9]{VP}.

@@

 Set $\mathcal{R}^\Phi_{n+1} = \mathcal{R}^\Phi_{\leq n+1} \setminus \mathcal{R}^\Phi_{\leq n}$.
 We need to verify that the image of every relator $R \in \mathcal{R}^\Phi_{n+1}$ under $\widetilde{j}^\Phi_n$ is a trivial element of $\St^{\leq n}(\Phi, B)$. In the special case $\Phi=\rA_{\geq 4}$ this assertion has already been demonstrated by Tulenbayev in~\cite[Lemma~3.3]{Tu83}. We will deduce the assertion in the remaining cases $\Phi=\rD_\ell,\rE_\ell$ from this Tulebayev's result.

 Fix a relator $R \in \mathcal{R}^\Phi_{n+1}$. By~\cref{superfluous-relations} we may assume that $R$ is not of type $\text{R3}^\bot$, hence the roots $\alpha, \beta$ appearing in $R$ are contained in a root subsystem of $\Phi$ of type $\rA_2$. Our assumptions on $\Phi$ guarantee that there exists some root subsystem $\Psi$ of type $\rA_4$ containing $\alpha$ and $\beta$. Consider the following commutative diagram in which the vertical arrows are induced by the embedding $\Psi\subseteq\Phi$. 
  \begin{equation} \nonumber \xymatrix{
 F(\mathcal{X}_{\leq n+1}^\Psi) \ar[r]^{j_n^\Psi} \ar[d] & \St^{\leq n}(\Psi, B) \ar[d] \\
 F(\mathcal{X}_{\leq n+1}^\Phi) \ar[r]^{j_n^\Phi} & \St^{\leq n}(\Phi, B) }
 \end{equation}
The relation $R$ lies in the image of the left arrow, therefore it comes from some relation $R' \in \mathcal{R}^\Psi_{n+1}$. The image of $R'$ in $\St^{\leq n}(\Psi, B)$ under $j_n^\Psi$ is trivial by Tulenbayev's result. But this implies that the image of $R$ under $\widetilde{j}_n^\Phi$ is also trivial and hence that $\widetilde{j}_n^\Phi$ gives rise to the desired map $j_n$.
\end{proof}

>>>>

\subsection{Steinberg groups}
Let $\Phi$ be a reduced and irreducible root system of rank $\geq 2$ and $R$ be a commutative ring with $1$. Recall that in this case the \emph{Steinberg group} $\St(\Phi, R)$ can be defined by means of generators $x_{\alpha}(s)$ (where $\alpha\in\Phi, s\in R$) and relations:
\begin{align}
& \phantom{[}
x_\alpha(s) \cdot x_\alpha(t) = x_\alpha(s+t),\ \alpha\in\Phi,\ s,t\in R; \label{rel:add}\\
& [x_\alpha(s), x_\beta(t)] = \prod\limits_{i,j\in\mathbb{N}}
 x_{i\alpha + j\beta}\left(N_{\alpha,\beta ij}\, s^i t^j\right),\quad \alpha,\beta\in\Phi,\ \alpha\neq\pm\beta,\ s,t\in R. \label{rel:CCF}
\end{align}
The indices $i$, $j$ appearing in the right-hand side of the above relation range over
all positive natural numbers such that $i\alpha + j\beta\in\Phi$.
The constants $N_{\alpha \beta i j}$ appearing in the right-hand side of \eqref{rel:CCF} are integers equal to $\pm 1,\pm 2,\pm 3$, they are called the {\it structure constants} of the Chevalley group $G_{sc}(\Phi, R)$. Several different methods have been proposed for choosing signs of these constants, see e.\,g.~\cite{V00}, \cite[\S~9]{VP}.

@@

 Set $\mathcal{R}^\Phi_{n+1} = \mathcal{R}^\Phi_{\leq n+1} \setminus \mathcal{R}^\Phi_{\leq n}$. It suffices to verify that the image of every relator $R \in \mathcal{R}^\Phi_{n+1}$ under $\widetilde{j}^\Phi_n$ is a trivial element of $\St^{\leq n}(\Phi, B)$. In the special case $\Phi=\rA_{\geq 4}$ this has already been demonstrated by Tulenbayev in~\cite[Lemma~3.3]{Tu83}, so in this case the proof of the proposition is complete. We will deduce the assertion in the remaining cases $\Phi=\rD_\ell,\rE_\ell$ from the special case $\Phi=\rA_4$ of Tulenbayev's result.

 Let $R$ be a relation from $\mathcal{R}^\Phi_{n+1}$. By~\cref{superfluous-relations} we may assume that $R$ is not of type $\text{R3}^\bot$, hence the roots $\alpha, \beta$ appearing in $R$ are contained in a root subsystem of $\Phi$ of type $\rA_2$. Our assumptions on $\Phi$ guarantee that there exists some root subsystem $\Psi$ of type $\rA_4$ containing $\alpha$ and $\beta$. Consider the following commutative diagram in which the vertical arrows are induced by the embedding $\Psi\subseteq\Phi$. 
  \begin{equation} \nonumber \xymatrix{
 F(\mathcal{X}_{\leq n+1}^\Psi) \ar[r]^{j_n^\Psi} \ar[d] & \St^{\leq n}(\Psi, B) \ar[d] \\
 F(\mathcal{X}_{\leq n+1}^\Phi) \ar[r]^{j_n^\Phi} & \St^{\leq n}(\Phi, B) }
 \end{equation}
The relation $R$ lies in the image of the left arrow, therefore it comes from some relation $R' \in \mathcal{R}^\Psi_{n+1}$. The image of $R'$ in $\St^{\leq n}(\Psi, B)$ under $j_n^\Psi$ is trivial by Tulenbayev's result. But this implies that the image of $R$ under $\widetilde{j}_n^\Phi$ is also trivial and hence that $\widetilde{j}_n^\Phi$ gives rise to the desired map $j_n$.
\end{proof}

\begin{rem}
 We needed to exclude relations of type $\text{R3}^\bot$ from the presentation of $\St^{\leq n}(\Phi, B)$ in the proof of the above proposition because already in the case $\Phi=\rD_\ell$ the pair of simple roots $\alpha_{\ell-1}$, $\alpha_{\ell}$ cannot be embedded into a root subsystem of type $\rA_4$.

 In the special case $M=A$, $B = A[t, t\inv]$ the assertion of the above proposition also holds in the cases $\Phi=\rA_2, \rA_3, \rD_4$. This is an easy consequence of the presentation of D.~Allcock applied to the affine Steinberg group $\St(\Phi, A[t, t\inv]) \cong \St(\widetilde{\Phi}^{(1)}, A)$ (Allcock's presentation implies that $\St(\Phi, A[t, t\inv])$ can be presented using only generators and relations of degree $\leq 1$ with respect to both $t$ and $t^{-1}$, see~\cite[Corollary~1.3]{A13}).

 In the cases $\Phi = \rA_3, \rD_4$, $M \neq A$ it is still possible to prove the injectivity of $i_n$ starting from $n\geq 2$ using a variation of the argument of Rehmann--Soul{\'e} (cf. the lower bound for $m$ in~\cite[3.2.1]{RS76}). However, apparently, it is not possible to establish the injectivity of $i_1$ in the specified cases using arguments similar to~\cite{RS76}.
\end{rem}
